---
title: "Assignment 1 DATA607"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mario Pena
## 9/1/2019

A dataset was loaded from the internet onto a data frame from the **Mushrooms** dataset and the following code was implemented: 


```{r}
mushrooms<-read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data", header= FALSE, sep=",")
```

We can see the dimensions of the data. This command should output [1] 8124 23, indicating that there are 8124 rows and 23 columns:

```{r}
dim(mushrooms)
```

The column names were changed to reflect their full description:

```{r}
colnames(mushrooms) <-c("EDIBILITY","CAP-SHAPE","CAP-SURFACE","CAP-COLOR","BRUISES","ODOR","GILL-ATTACHMENT","GILL-SPACING","GILL-SIZE","GILL-COLOR","STALK-SHAPE","STAL-ROOT","STAL-SURFACE-ABOVE-RING", "STALK-SURFACE-BELOW-RING", "STALK-COLOR-ABOVE-RING", "STALK-COLOR-BELOW-RING", "VEIL-TYPE", "VEILCOLOR", "RING-NUMBER", "RING-TYPE", "SPORE-PRINT-COLOR", "POPULATION", "HABITAT")
```

And now we can see the names:

```{r}
names(mushrooms)
```

Some of the characters representing data in the rows were changed to their actual name for a few variables, which will be used in the subset below:

```{r}
mushrooms$EDIBILITY <- sub("p", "Poisonous", mushrooms$EDIBILITY) 
mushrooms$ODOR <- sub("m", "Musty", mushrooms$ODOR) 
mushrooms$POPULATION <- sub("c", "Clustered", mushrooms$POPULATION) 
mushrooms$HABITAT <- sub("d", "Woods", mushrooms$HABITAT)
```

A subset of the columns in the dataset was created:

```{r}
subset(mushrooms, EDIBILITY == "Poisonous" & ODOR == "Musty", select = c(EDIBILITY, ODOR, POPULATION, HABITAT))
```

What is the mushrooms’ most common veil color? In this case a barplot is used to find the answer:

```{r echo=FALSE}
mushrooms$VEILCOLOR <- sub("n", "Brown", mushrooms$VEILCOLOR)
mushrooms$VEILCOLOR <- sub("^o$", "Orange", mushrooms$VEILCOLOR)
mushrooms$VEILCOLOR <- sub("^w$", "White", mushrooms$VEILCOLOR)
mushrooms$VEILCOLOR <- sub("y", "Yellow", mushrooms$VEILCOLOR)
```

```{r}
barplot(table(mushrooms$VEILCOLOR), ylim=c(0,8000), ylab="Frequency", main="Barplot of Veil Color Distribution")
```

“White” is the most common veil color by far.
